Day 1. Basic R Manipulation
=============================

# Basic data format

## {.tabset}

### task

Constuct a numeric vector from 1 to 10.

### solution

```{r}
x = 1:10
x
```

## {.tabset}

### task

Constuct a numeric vector of 3, 6, 2, 9, 1.

### solution

```{r}
x = c(3, 6, 2, 9, 1)
x
```


## {.tabset}

### task

Constuct a character vector of weekdays in a week, then extract the weekdays for weekend.

### solution

```{r}
x = c("Monday", "Tuesday", "Wendesday", "Thursday", "Friday", "Saturday", "Sunday")
x
x[6:7]
```

## {.tabset}

### task

Constuct a numeric vector of when you wake up in the morning in a week and use weekday as 
the name of the vector, then extract two days of weekdend.

### solution

```{r}
# way 1:
x = c("Monday" = 7, "Tuesday" = 7, "Wendesday" = 8, "Thursday" = 9, "Friday" = 8, "Saturday" = 12, "Sunday" = 12)
x
# way 2:
x = c(7, 7, 8, 9, 8, 12, 12)
names(x) = c("Monday", "Tuesday", "Wendesday", "Thursday", "Friday", "Saturday", "Sunday")
x

x[6:7] # or
x[c("Saturday", "Sunday")]
```

## {.tabset}

### task

Construct a matrix of random values with 10 rows and 5 columns.

consider `rnorm()`, `sample()`, `runif()` for random value generation.

### solution

```{r}
m = matrix(rnorm(50), nrow = 10)
m = matrix(runif(50), nrow = 10)
m = matrix(sample(10, 10, replace = TRUE), nrow = 10)
```

## {.tabset}

### task

construct a data frame of monthly climate statistics in Heidelberg (https://en.wikipedia.org/wiki/Heidelberg#collapsibleTable0).

### solution

```{r}
climate = data.frame(
	average_high  = c(3.8, 6.1, 10.9, 15.4, 19.9, 23, 25.5, 25.1, 21.5, 15.3, 8.5, 4.8),
	daily_mean    = c(2.4, 3.7, 7.4, 11.2, 15.5, 18.1, 20.6, 20.1, 16.1, 11.5, 6.3, 3.3),
	average_low   = c(-1.4, -0.7, 1.9, 4.9, 8.9, 12.2, 14, 13.8, 10.6, 6.7, 2.4, -0.4),
	precipitation = c(56, 53, 53, 61, 79, 86, 71, 66, 53, 58, 66, 66),
	sunshine      = c(46, 78, 118, 173, 206, 215, 233, 219, 157, 101, 50, 35)
)
rownames(climate) = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
```

Note: we will use `climate` variable later.

## {.tabset}


### task

construct a list which contains dates in each month this year.

construct a list which contains some of the following description of yourself:

- name
- age
- interest (more than one)
- do you like programming

### solution

```{r}
date = list(
	Jan = 1:31, 
	Feb = 1:28, 
	Mar = 1:31, 
	Apr = 1:30, 
	May = 1:31, 
	Jun = 1:30,
	Jul = 1:31,
	Aug = 1:31, 
	Sep = 1:30,
	Oct = 1:31,
	Nov = 1:30,
	Dec = 1:31
)

me = list(
	name = "Zuguang Gu",
	age = 34,
	interest = c("movie", "game"),
	like_programming = TRUE
)
me
me$name
```


# read/write 

## {.tabset}

### task

save `climate` as a csv file

save `climate` as a tab-separated file

### solution

```{r}
write.csv(climate, file = "climate.csv")
write.table(climate, sep = "\t", quote = FALSE, file = "climate.tab")
```

## {.tabset}

### task

read climate data back to R

### solution

```{r}
df = read.csv("climate.csv", header = TRUE, row.names = 1)
df = read.table("climate.tab", header = TRUE, row.names = 1)
```

# Descriptive statistics

## {.tabset}

### task

With `climate`, calcualte the mean/median/sd/ of each column:

### solution

```{r}
mean(climate$average_high)
mean(climate$daily_mean)
mean(climate$precipitation)

apply(climate, 2, mean)
apply(climate, 2, median)
apply(climate, 2, sd)
```


## {.tabset}

### task

Does temporature correlates to the precipitation or sunshine?

### solution

```{r}
cor(climate$daily_mean, climate$precipitation)
cor(climate$daily_mean, climate$sunshine)
```

## {.tabset}

### task

Does the precipitation in May, June, July, Octorber, September and Octorber
has difference to to other monthe?

### solution

```{r}
precipitation = climate$precipitation
t.test(precipitation[5:10],
	   precipitation[c(1:4, 11:12)])

```
