Day 1. Basic R Manipulation
=============================

# Basic data format

## {.tabset}

### task

Constuct a numeric vector from 1 to 10.

### solution

```{r}
x = 1:10
x
```

## {.tabset}

### task

Constuct a numeric vector of 3, 6, 2, 9, 1.

### solution

```{r}
x = c(3, 6, 2, 9, 1)
x
```


## {.tabset}

### task

Constuct a character vector of weekdays in a week, then extract the weekdays for weekend.

### solution

```{r}
x = c("Monday", "Tuesday", "Wendesday", "Thursday", "Friday", "Saturday", "Sunday")
x
x[6:7]
```

## {.tabset}

### task

Constuct a numeric vector of when you wake up in the morning in a week and use weekday as 
the name of the vector, then extract two days of weekdend.

### solution

```{r}
# way 1:
x = c("Monday" = 7, "Tuesday" = 7, "Wendesday" = 8, "Thursday" = 9, "Friday" = 8, "Saturday" = 12, "Sunday" = 12)
x
# way 2:
x = c(7, 7, 8, 9, 8, 12, 12)
names(x) = c("Monday", "Tuesday", "Wendesday", "Thursday", "Friday", "Saturday", "Sunday")
x

x[6:7] # or
x[c("Saturday", "Sunday")]
```

## {.tabset}

### task

Construct a matrix of random values with 10 rows and 5 columns.

consider `rnorm()`, `sample()`, `runif()` for random value generation.

### solution

```{r}
m = matrix(rnorm(50), nrow = 10)
m = matrix(runif(50), nrow = 10)
m = matrix(sample(10, 10, repleace = TRUE), nrow = 10)
```

## {.tabset}

### task


